---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: "Lian Shen"
date: "February 26, 2019"
output: 
  html_document:
    self_contained: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = 'center')
```

```{css, echo=FALSE}
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
```


***

Github repository: https://github.com/lianshen/compscix-415-2-assignments 

Load libraries
```{r load_libraries}
library(tidyverse)
```

***
##**The tidyverse packages**
###**1.**  
**Can you name which package is associated with each task below?**  
**• Plotting -** `ggplot2`  
**• Data munging/wrangling -** `dplyr`  
**• Reshaping (speading and gathering) data -** `tidyr`  
**• Importing/exporting data -** `readr`  

***
###**2.**  
**Now can you name two functions that you’ve used from each package that you listed above for these tasks?**  
**• Plotting -** `geom_bar()` and `geom_histogram()`  
**• Data munging/wrangling -** `filter()` and `summarize()`  
**• Reshaping data -** `gather()` and `spread()`  
**• Importing/exporting data -** `read_csv()` and `write_delim()`  

***
##**R Basics**  
###**1.**  
**Fix this code *with the fewest number of changes possible* so it works:**  
**`My_data.name___is.too00ooLong! <- c( 1 , 2 , 3 )`**  
```{r}
#remove invalid ! character
My_data.name___is.too00ooLong <- c( 1 , 2 , 3 )
```

***
###**2.**  
**Fix this code so it works:**  
**`my_string <- C('has', 'an', 'error', 'in', 'it)`**
```{r}
#combine function c() is lowercase. must end last string argument with quote '
my_string <- c('has', 'an', 'error', 'in', 'it')
```

***
###**3.**
**Look at the code below and comment on what happened to the values in the vector.**  
**`my_vector <- c(1, 2, '3', '4', 5)`**  
**`my_vector`**  

**`## [1] "1" "2" "3" "4" "5"`**  

All values in the vector were converted to characters because  
1. `c()` combines its arguments and converts them to a common type while doing so.  
2. The common type decided upon depends on the highest type of argument listed following the hierarchy: `NULL` < `raw` < `logical` < `integer` < `double` < `complex` < `character` < `list` < `expression`. Out of the types listed as arguments (`character` and `integer`), `character` types are higher than `integer` types.

***
##**Data import/export**
###**1.**  
**Download the rail_trail.txt file from Canvas (in the Midterm Exam section) and successfully import it into R. Prove that it was imported successfully by including your import code and taking a `glimpse` of the result.**  

```{r}
filepath <- "C:\\Users\\lshen\\Documents\\Data Science\\rail_trail.txt"
#rail_trail.txt file is "|" delimited
rail_trail_data <- read_delim(file = filepath, delim = '|')
glimpse(rail_trail_data)
```

***
###**2.**  
**Export the file into a comma-separated file and name it “rail_trail.csv”. Make sure you define the `path` correctly so that you know where it gets saved. Then reload the file. Include your export and import code and take another `glimpse`.**  

```{r}
new_filepath <- "C:\\Users\\lshen\\Documents\\Data Science\\rail_trail.csv"
rail_trail_data %>% write_csv(path = new_filepath)
new_rail_trail_data <- read_csv(file = new_filepath)
glimpse(new_rail_trail_data)
```

***
##**Visualization**
###**1.**  
**Critique this graphic: give only three examples of what is wrong with this graphic. Be concise.** 
1. Using circles (area) as a metric for response count is difficult to compare.  
2. The categories used is not consistent: the graphic is first comparing by age group, then by gender.  
3. There is no point in coloring by gender if the age group classes don't specify gender ratio.

***
###**2.** 
**Reproduce this graphic using the diamonds data set.**  

```{r}
diamonds %>% ggplot(aes(x = cut, y = carat, fill=color)) + geom_boxplot(position = "identity") + xlab("CUT OF DIAMOND") + ylab("CARAT OF DIAMOND") + coord_flip()
```

***
###**3.**
**The previous graphic is not very useful. We can make it much more useful by changing one thing about it. Make the change and plot it again.**

```{r}
diamonds %>% ggplot(aes(x = cut, y = carat, fill=color)) + geom_boxplot(position = "dodge") + xlab("CUT OF DIAMOND") + ylab("CARAT OF DIAMOND") + coord_flip()
```

***
##**Data munging and wrangling**
###**1.**
**Is this data “tidy”? If yes, leave it alone and go to the next problem. If no, make it tidy. *Note: this data set is called table2 and is available in the tidyverse package. It should be ready for you to use after you’ve loaded the tidyverse package*.**

```{r}
table2
```

```{r}
#tidy table2
table2 %>% spread(key = type, value = count)
```


***
###**2.**
**Create a new column in the diamonds data set called `price_per_carat` that shows the price of each diamond per carat (hint: divide). Only show me the code, not the output.**

```{r, results='hide'}
diamonds %>% mutate(price_per_carat = price/carat)
```


***
###**3.**
**For each cut of diamond in the diamonds data set, how many diamonds, and what proportion, have a price > 10000 and a carat < 1.5? There are several ways to get to an answer, but your solution must use the data wrangling verbs from the tidyverse in order to get credit.**  
```{r}
expensive_diamonds <- diamonds %>% select(cut, price, carat) %>% group_by(cut) %>% mutate(total_cut = n()) %>% filter(price > 10000 & carat < 1.5) %>% mutate(expensive_count = n()) %>% mutate(proportion = expensive_count/total_cut) %>% select(cut, expensive_count, total_cut, proportion) %>% distinct()
View(expensive_diamonds)
expensive_diamonds %>% ggplot(aes(x=cut, y=proportion)) + geom_col()
```

**• Do the results make sense? Why?**  
**• Do we need to be wary of any of these numbers? Why?**  

***
##**EDA**  
**Take a look at the `txhousing` data set that is included with the `ggplot2` package and answer these questions:**


###**1.**  
**During what time period is this data from?**

***
###**2.**
**How many cities are represented?**

***
###**3.**
**Which city, month and year had the highest number of sales?**

***
###**4.**
**What kind of relationship do you think exists between the number of listings and the number of sales? Check your assumption and show your work.**

***
###**5.**
**What proportion of sales is missing for each city?**

***
###**6.**
**Looking at only the cities and months with greater than 500 sales:**  
**• Are the distributions of the median sales price (column name median), when grouped by city, different? The same? Show your work.**  
**• Any cities that stand out that you’d want to investigate further?**  
**• Why might we want to filter out all cities and months with sales less than 500?**  

***
Once you are finished with your midterm, commit your final changes with the comment “finished the midterm
- woohoo” and push your R Markdown file and your html or pdf file to Github.