---
title: "COMPSCIX 415.2 Homework 3"
author: "Lian Shen"
date: "February 12, 2019"
output: 
  html_document:
    self_contained: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = 'center')
library(tidyverse)
```
***
##**3.8.1 Exercises**
####**1. What is the problem with this plot? How could you improve it?**
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + geom_point()
```


This plot only shows a subset of the datapoints because of the discrete nature of the dataset. There are datapoints overlapping with each other. The code below should show more datapoints.

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + geom_point(position = "jitter")
```

***
####**2. What parameters to `geom_jitter()` control the amount of jittering?**

The `width` and `height` parameters adjust the horizontal and vertical amount of jitter/randomness, respectively.

***
####**3. Compare and contrast `geom_jitter()` with `geom_count()`.**

```{r}
mpgPlot <- ggplot(data = mpg, mapping = aes(x = cty, y = hwy))
mpgPlot + geom_jitter()
mpgPlot + geom_count()
```

`geom_jitter()` is a shortcut for `geom_point(position = "jitter")` for adding randomization to the location of each point. `geom_count()` is an extension to `geom_point()` where each point's area reflects the number of observations of the location. Both geoms are useful for discretizing data where overlapping points may be an issue.

***
####**4. What’s the default position adjustment for `geom_boxplot()`? Create a visualisation of the `mpg` dataset that demonstrates it.**

The default position parameter is "dodge2". "Dodge2"" adjusts the horizontal position of the boxplot to prevent the sides of the plots from overlapping.
```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + geom_boxplot()
```

***
##**3.9.1 Exercises**
####**2. What does `labs()` do? Read the documentation.**
`labs()` modifies the axes, legend, and plot labels. It provides the option to set the title, subtitle, caption, tags, and other labels.

***
####**4. What does the plot below tell you about the relationship between city and highway `mpg`? Why is `coord_fixed()` important? What does `geom_abline()` do?**
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + geom_point() + geom_abline() + coord_fixed()
```
The plot shows a linear trend between city and highway. `coord_fixed()` fixes the cartesion coordinate plot's aspect ratio to 1 by default so that one unit of the x-axis is the same as one unit on the y-axis. `coord_fixed()` is important for adjusting the aspect ratio to whatever is desired and is a good way to keep individual plots' aspect ratios consistent with each other.  
`geom_abline()` plots a diagonal reference line where the intercept and slope can be set. 

***
##**4.4 Practice**
####**1. Why does this code not work?**
**`my_variable <- 10`**  
**`my_varıable`**  
**`#> Error in eval(expr, envir, enclos): object 'my_varıable' not found`**  

`my_varıable` does not exist and has no value assigned to it. One character (ı) is off and should be 'i'. The variable that was set is called `my_variable`.

***
####**2. Tweak each of the following R commands so that they run correctly:**

**`library(tidyverse)`**  

**`ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))`**  

**`fliter(mpg, cyl = 8)`**  
**`filter(diamond, carat > 3)`**  

`fliter(mpg, cyl = 8)` should be `filter(mpg, cyl == 8)`.  
`filter(diamond, carat > 3)` should be `filter(diamonds, carat > 3)`.

***
##**5.2.4 Exercises**
####**1. Find all flights that**  

```{r}
library(nycflights13)
```

*1. Had an arrival delay of two or more hours*   
```{r}
flights %>% filter(arr_delay >= 120)
```

*2. Flew to Houston (IAH or HOU)*  
```{r}
flights %>% filter(dest %in% c('IAH','HOU'))
```

*3. Were operated by United, American, or Delta*  
```{r}
airlines
flights %>% filter(carrier %in% c('AA','DL','UA'))
```

*4. Departed in summer (July, August, and September)*  
```{r}
flights %>% filter(month %in% c(7,8,9))
```

*5. Arrived more than two hours late, but didn’t leave late*  
```{r}
flights %>% filter(arr_delay > 120 & dep_delay <= 0)
```

*6. Were delayed by at least an hour, but made up over 30 minutes in flight*  

If "made up over 30 minutes in flight" means saved over 30 minutes from the scheduled flight time:  
```{r}
mutatedFlights <- flights %>% mutate(sched_flight_duration = sched_arr_time - sched_dep_time, actual_flight_duration = arr_time - dep_time)
mutatedFlights <- mutatedFlights %>% mutate(made_up_time = sched_flight_duration - actual_flight_duration)
mutatedFlights %>% filter(dep_delay >= 60 & made_up_time > 30)
```

If "made up over 30 minutes in flight" means the arrival delay was more than 30 minutes less than the departure delay:  
```{r}
flights %>% filter(dep_delay >= 60 & arr_delay < dep_delay - 30)
```


*7. Departed between midnight and 6am (inclusive)*  
```{r}
flights %>% filter(dep_time >= 0 & dep_time <= 600)
```

***
####**3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?**

```{r}
flights %>% filter(is.na(dep_time))
```
There are 8,255 flights that are missing depature times. Of these records, `dep_delay`, `arr_time`, `arr_delay`, and `air_time` are also missing values. These rows may represent cancelled flights that never departed which would explain why the other variables are missing values as well.

***
####**4. Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)**

`NA ^ 0` returns 1 because `NA` acts as a placeholder number (could be `Inf`, `0`,). Replacing `NA` with any number (including `Inf`) and raising it to the power of 0 will always return 1. However, `NA * 0` returns `NA` because there are cases where a number multiplied by 0 is indeterminate. Since `NA` could represent `Inf`, `Inf * 0` is indeterminate, so `NA * 0` will return `NA`.  
`NA | TRUE` returns `TRUE` because the logical *OR* operator will return `TRUE` if any argument is `TRUE`. However, if `NA | FALSE` was executed, the result would be `NA` because `NA` could represent `TRUE` or `FALSE`. If `NA` represented `FALSE`, the expression would return `FALSE`, but if `NA` represented `TRUE`, the expression would return `TRUE`. Therefore, the expression is indeterminate and returns `NA`.  
`NA & FALSE` returns `FALSE` because the logical *AND* operator will return `FALSE` if any argument is `FALSE`. `NA` could represent `TRUE` or `FALSE`, but in both scenarios, the expression will always return `FALSE`.  
The general rule is that if a consistent result can be obtained regardless of whatever value `NA` could represent, the expression will yield that consistent result. If the expression cannot yield a consistent result for what `NA` could represent, the expression will return `NA`.

***
##**5.4.1 Exercises**
####**1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.**

```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay) %>% variable.names()
flights %>% select(dep_time, dep_delay, arr_time, arr_delay) %>% variable.names()
cols <- c(dep_time = "dep_time", dep_delay = "dep_delay", arr_time = "arr_time", arr_delay = "arr_delay")
select(flights, !!cols) %>% variable.names()
flights %>% select (!!cols) %>% variable.names()
select(flights, starts_with("dep"), starts_with("arr")) %>% variable.names()
flights %>% select(starts_with("dep"), starts_with("arr")) %>% variable.names()
select(flights, dep_time:arr_delay, -sched_dep_time, -sched_arr_time) %>% variable.names()
flights %>% select(dep_time:arr_delay, -sched_dep_time, -sched_arr_time) %>% variable.names()
select(flights, ends_with("delay"), ends_with("time"), -sched_dep_time, -sched_arr_time, -air_time) %>% variable.names()
flights %>% select(ends_with("delay"), ends_with("time"), -sched_dep_time, -sched_arr_time, -air_time) %>% variable.names()
vars <- c("dep_time", "dep_delay", "arr_time", "arr_delay")
select(flights, one_of(vars)) %>% variable.names()
flights %>% select(one_of(vars)) %>% variable.names()
```

`variable.names()` is called for brevity to show that the `select` function returns the correct columns, assuming the data is correct.

***
####**3. What does the one_of() function do? Why might it be helpful in conjunction with this vector?**
**`vars <- c("year", "month", "day", "dep_delay", "arr_delay")`**

`one_of()` is a helper function for `select` that matches the data's variable names in a character vector argument. If the `select` operation uses `one_of` with the vector, the columns/variables returned are those that match any of the characters of the vector.

***



