---
title: "COMPSCIX 415.2 Homework 1"
output: html_document
author: "Lian Shen"
date: 2/1/2019
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
***
My Github repository for my assignments can be found at this URL: https://github.com/lianshen/compscix-415-2-assignments  

***
```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
data("WorldCities")
```
***
**Question 1**: There are 23,018 observations and 10 variables. Some of the variables are called name, latitude, longitude, counter, and population.

```{r world_cities}
glimpse(WorldCities)
WorldCities <- head(WorldCities, 200) # 200 rows
country_col <- WorldCities$country
unique(country_col)
```
***
**Question 2**: There is a column called region in the data set. Can you extract this and show
only the unique values?

```{r region_column}
region_col <- WorldCities$region
unique(region_col)
```
***
**Question 3**: Can you extract and show the unique entries from the country column in
WorldCities using one line of code and two %>% operators?

```{r unique_country}
WorldCities %>% select(country) %>% unique()
```
***
Visualization Demo  
```{r demo_visualize}
WorldCities %>% ggplot(aes(x = population)) + geom_histogram()
```

***
**Question 4**: Make one more histogram of the population and add a new x-label, a new ylabel,
a new title (use ggtitle('my title')), and change the theme of the plot using theme_bw().

```{r world_cities_visualization}
WorldCities %>% ggplot(aes(x = population)) + geom_histogram() + xlab('City Population') + ylab('Number of Cities') + ggtitle('Population of World Cities Histogram') + theme_bw()
```