---
title: "COMPSCIX 415.2 Homework 4"
author: "Lian Shen"
date: "February 19, 2019"
output: 
  html_document:
    self_contained: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
***
##**5.6.7 Exercises**
####**2. Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).**

***
####**4. Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?**

***
####**6. What does the sort argument to count() do. When might you use it?**

***
####**5. Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))**

***
##**10.5 Exercises**
####**1. How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).**

***
####**2. Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?**  

**`df <- data.frame(abc = 1, xyz = "a")`**  
**`df$x`**  
**`df[, "xyz"]`**  
**`df[, c("abc", "xyz")]`**  

***
####**3. If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?**

***
####**6. What option controls how many additional column names are printed at the footer of a tibble?**

***
##**12.3.3 Exercises**
####**2. Why does this code fail?**  

**`table4a %>% gather(1999, 2000, key = "year", value = "cases")`**  
**`#> Error in inds_combine(.vars, ind_list): Position must be between 0 and n`**  

***
####**3. Why does spreading this tibble fail? How could you add a new column to fix the problem?**  

```{r}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
) 
```

***
####**4. Tidy the simple tibble below. Do you need to spread or gather it? What are the variables?**  
```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)
```

***
##**12.4.3 Exercises**
####**1. What do the extra and fill arguments do in separate()? Experiment with the various options for the following two toy datasets.**  
```{r}

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))
```
***
####**2. Both unite() and separate() have a remove argument. What does it do? Why would you set it to FALSE?**  

***
####**Follow these steps:**  
**- Download the baby_names.txt file from Canvas which is in the Homework 4 assignment section.**  
**- Load this file into R correctly and take a glimpse of the output.**  
**- Export this file as a csv file and call it ‘baby_names.csv’.**  
**- Reload the baby_names.csv file and take another glimpse.**  
**- Show all of your code and the output. There should be two data import lines of code, one data export line of code, and two glimpses of the data.**  
